---
output: pdf_document
---

Packages: 


```{r}
library(dplyr)
```






Upload SPI Time Series Index into R: 



```{r}
library(readxl)
SPI_Time_Series_1990_2020 <- Copy_of_SPI_Time_Series_1990_2020_data_file_to_take_notes <- read_excel("/Users/justus/Desktop/Yale University Stuff/Tobin Research Assistantship/Copy of SPI Time Series 1990-2020 data file to take notes.xlsx", 
 sheet = "1990-2020 Time-series data")


```


Get rid of all columns after the specific indicators that make-up the basic
human-wellbing dimension: 



```{r}

Basic_Human_Needs <- SPI_Time_Series_1990_2020[, 1:42]





```



Get rid of irrelevant indicators. 


```{r}

Basic_Human_Needs <- Basic_Human_Needs[, -c(9, 10, 11, 16:24)]



```



Adjust column names: 


```{r}


colnames(Basic_Human_Needs) <- Basic_Human_Needs[1, ]
Basic_Human_Needs <- Basic_Human_Needs[-1, ]




```


Open daily caloric supply of capita: 


```{r}

daily_per_capita_caloric_supply <- read.csv("daily-per-capita-caloric-supply.csv")


#Exclude Continents and regions: 

daily_per_capita_caloric_supply <- daily_per_capita_caloric_supply %>% filter(Code != "")


#adjusting to time series of 1990-2020

daily_per_capita_caloric_supply <- daily_per_capita_caloric_supply[daily_per_capita_caloric_supply$Year >= 1990 & daily_per_capita_caloric_supply$Year <= 2020, ]


```



Add daily calorie supply to SPI time series: 


```{r}

# Rename columns for merging
names(daily_per_capita_caloric_supply)[names(daily_per_capita_caloric_supply) == "Entity"] <- "Country"
names(daily_per_capita_caloric_supply)[names(daily_per_capita_caloric_supply) == "Year"] <- "SPI \r\nyear"


# Merge datasets by 'Country' and 'SPI year'
Basic_Human_Needs <- merge(Basic_Human_Needs, daily_per_capita_caloric_supply[, c("Country", "SPI \r\nyear", "Daily.calorie.supply.per.person")],
                     by = c("Country", "SPI \r\nyear"),
                     all.x = TRUE)



```




Open Per Capita Protein supply from all foods per day: 



```{r}

daily_protein_supply_per_person <- read.csv("/Users/justus/Desktop/Yale University Stuff/Tobin Research Assistantship/Basic Human Needs Indicator /protein_supply.csv")

```



Getting rid of irrelevant columns: 


```{r}

daily_protein_supply_per_person <- daily_protein_supply_per_person[, c(2, 3, 39)]



```



add daily protein supply to index: 


```{r}
# Rename columns for merging
names(daily_protein_supply_per_person)[names(daily_protein_supply_per_person) == "Country"] <- "Country"
names(daily_protein_supply_per_person)[names(daily_protein_supply_per_person) == "Year"] <- "SPI \r\nyear"


# Merge datasets by 'Country' and 'SPI year'
Basic_Human_Needs <- merge(Basic_Human_Needs, daily_protein_supply_per_person[, c("Country", "SPI \r\nyear", "Food.supply..Protein.g.per.capita.per.day.")],
                     by = c("Country", "SPI \r\nyear"),
                     all.x = TRUE)

```





Open Health Life Expectancy (HALE) at birth: 


```{r}
life_expectancy_at_birth <- read.csv("/Users/justus/Desktop/Yale University Stuff/Tobin Research Assistantship/Basic Human Needs Indicator /Health Life Expectance (HALE) at birth.csv")

```

Get rid of irrelevant columns/variables: 

```{r}

life_expectancy_at_birth <- subset(life_expectancy_at_birth, Indicator == "Healthy life expectancy (HALE) at birth (years)" & Dim1 == "Both sexes")
life_expectancy_at_birth <- life_expectancy_at_birth[, c(8, 10, 24)]



```


add life expectancy at birth to index: 


```{r}
# Rename columns for merging
names(life_expectancy_at_birth)[names(life_expectancy_at_birth) == "Location"] <- "Country"
names(life_expectancy_at_birth)[names(life_expectancy_at_birth) == "Period"] <- "SPI \r\nyear"


# Merge datasets by 'Country' and 'SPI year'
Basic_Human_Needs <- merge(Basic_Human_Needs, life_expectancy_at_birth[, c("Country", "SPI \r\nyear", "FactValueNumeric")],
                     by = c("Country", "SPI \r\nyear"),
                     all.x = TRUE)




```



































